Identification of high-yielding soybean lines with exceptional seed composition qualities

### 2022-06-30 doc/manuscript/manuscript.tex
Added a a step to calculate protein meal and re-ran the pipeline. lsmeans seem to have been calculated without issue.
Also added protein meal to the phenotype name conversion utility tables.

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{In current markets, the primary uses for soybean seed are in
products derived from it's oil or protein content. However, growers are
compensated based on seed oil so a more valuable crop is one with both
high yield, and high protein, oil, or a combination of the two. A
negative correlation between seed yield and seed protein makes improving
these traits simultaneously difficult but not impossible through
conventional breeding. A selection of lines with exceptional yield and
seed composition qualities was made from two recombinant inbred line
(RIL) soybean mapping populations to identify soybean varieties with
yield comparable to existing cultivars, and seed composition traits
superior to existing cultivars. The performance of these RILs were
evaluated in multiple environments and several genotypes were identified
with yield comparable to that of existing high-performance cultivars
with seed protein and or oil that is equal to or superior to the
high-performing cultivars. These genotypes will provide breeders with
additional sources of germplasm for continuing efforts to improve seed
composition traits, and provide growers with valuable genotype options.}

@@

Seed yield, oil, and protein are all valuable traits in a soybean
variety, however breeding lines which have both high yield and protein
has been difficult to develop due to the negative correlation between
the two
traits\citep{burton1987quantitative, ProtOilCorr, ProtOilCorr_new}.
While considerable efforts have been made to identify loci which control
these seed quality traits so that MAS breeding strategies can be
utilized for their improvement, to date the applications of such markers
have been few. This is largely due to the lack of markers which are
uniquely associated with one trait, and are also stable across genetic
and environmental backgrounds. While there is still reason to continue
this genetic research, it is important that breeders take every
opportunity to identify lines with both high yield, and seed composition
traits like oil and protein content so that new varieties can be
released.

Soybean lines typically contain about 20\% oil and 40\% protein content
on a dry weight basis\citep{ProteinGenomics}. The market for soybean
meal requires 47.5\% protein content in the meal, which corresponds to
approximately 41.5\% protein content on a dry weight
basis\citep{ProteinGenomics}. Oil and protein content are two of the
most important seed composition traits in soybean so if one is
decreased, the other should be correspondingly increased to account for
the loss in value. The inverse correlation between protein and oil
contents is well known and is suspected to be due at least partially to
the action of pleiotropic genes and competing metabolic pathways which
control the expression of each trait\citep{SeedCompositionGenomics}.

Despite the difficulty in simultaneously breeding for all three of these
traits, releases of varieties with elevated protein contents and only
moderately reduced yield has shown that it is not impossible. The high
protein germplasm lines R05-1415 and R05-1772 were released recently and
contain 46.9\% and 46.1\% protein content with while still producing
yields 94\% and 91\% of that of the high yielding 5002T
cultivar\citep{HiProtHiYield02}. Lines TN03-350 and TN04-5321 contain
43.9\% and 43.1\% protein content while having superior or comparable
performance to yield checks\citep{HiProtHiYield04}. The Prolina cultivar
has a protein content of 46.1\% with a yield only 13\% reduced from the
Centennial check\citep{HiProtHiYield01}. The Highpro1 cultivar was
released in 2016 and has a yield which is greater than or equal to 97\%
of that of the highest yielding check cultivar, IA3023 with a protein
content of 40.1\%\citep{HighPro1}. Cultivars produced through
conventional breeding techniques such as these have shown that it is
possible to identify lines with both high seed protein and seed yield.
Efforts to find these lines should be continued to provide growers and
breeders with additional high value cultivars, and germplasm which can
be further used to improve protein and yield traits.

To meet this goal, two recombinant inbred line (RIL) oil mapping
populations were screened for lines which showed promising combinations
of yield and seed composition traits. Successive rounds of selection
were conducted to identify and characterize lines with high values for
yield as well as protein and oil composition were performed between 2018
and 2021 to identify soybean lines with yield comparable to existing
check cultivars and protein and or oil composition superior to that of
the check cultivars.

@@

In 2018, oil mapping populations 201 and 202 were grown as plant rows at
the Central Crops Research Station in Clayton, NC. These populations
consisted of 273 and 237 recombinant inbred lines (RILs) respectively.
Several agronomic traits were scored in the field for each population.

@@

Following harvest, yield, seed weight, protein, and oil content were
measured after seed was air dried to approximately 7\% moisture content
in a greenhouse. Protein and oil contents were measured on a dry basis
using a Perten DA 7250 NIR®instrument. Yield and seed weight were
measured after seed had been sifted and cleaned of debris and cracked
seed.

To select lines for the 2019 growing season, lines with abnormally low
bulk weights or extreme maturity dates from 2018 were first removed from
consideration. Two yield trials were then developed for each mapping
population. The maturity dates of RILs were considered when forming
tests such that the lines of each test would have a maturity date range
approximately half that of the total mapping population from which it
was derived. RILs were selected for each test which were also
representative of the distribution of seed protein and seed oil traits
for each population.

Eighty unique lines were selected from each population which satisfied
these criteria, and each yield test was comprised of 40 RILs. Three
high-yielding check cultivars and the two parents of the respective
population were also included in each test. These yield tests were named
test 1 and test 2 for RILs derived from mapping population 201 and test
3 and 4 for RILs derived from mapping population 202. Yield check
cultivars Dunphy, Osage\citep{Osage}, and Roy were used in tests 1 and
2, while Dunphy, Dilday, and NC-Raleigh\citep{NCRaleighregistration}
were used for tests 3 and 4. These lines were selected to represent the
estimated maturities of the RILs in each test. The parents for tests 1
and 2 were cultivars LMN09-119 and N09-09, and the parents for tests 3
and 4 were LMN09-19 and N13-47.

@@

A strong negative correlation was observed in both tests between seed
protein and seed oil content. A coefficient of r = -0.93 was observed in
both populations. A negative correlation was observed between seed
protein and yield with correlation coefficient of r = -0.29 was observed
in both tests. This correlation was not statistically significant in
either population. A weak positive correlation was observed between seed
oil and seed yield in both populations. A correlation coefficient of r =
0.24 was observed in Test 1 while a coefficient of r = 0.28 was observed
in Test 2. These correlation coefficients were not statistically
significant in either population. Visualizations of the pairwise
correlation coefficients for Tests 1 and 2 can be seen in supplementary
figures 2 and 3, respectively.

\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{C:/Users/Jay/Desktop/Documents/R/Jay_yield/exports/plots/lsmean_scatterplot.pdf}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}

@@

Several RILs in each of the tests had yield that was comparable to that
of their check cultivars as per contrasts between RIL means and the
average of the check cultivars for each test. No genotypes had yield
that was quantitatively higher than that of the yield checks, however
many had comparable yield. Many of these genotypes with comparable yield
also had protein content that was superior to the check average in each
test. No genotypes had oil that was superior to the check cultivars, but
some had comparable yield and oil content as well as superior protein
content.

RILs were also evaluated for lodging and seed quality. All RILs had seed
quality that was on par with that of the high-yielding checks however,
however 11 genotypes had lodging scores that were statistically
significantly greater than the checks on the basis of contrasts. These
RILs were removed from consideration for recommendation on the basis of
their relatively poor agronomic performance.

>>>>

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{In current markets, the primary uses for soybean seed are in
products derived from it's oil or protein content. However, growers are
compensated based on seed yield so a more valuable crop is the one that
is high yielding with an optimum combination of protein and oil. A
negative correlation of seed protein with seed yield and oil makes
simultaneous improvement of these traits difficult but not impossible
through conventional breeding. A selection of lines with exceptional
yield and seed composition qualities was made from two recombinant
inbred line (RIL) soybean mapping populations to identify high protein
and/or high oil lines with yield comparable to elite cultivars. The
performance of these RILs were evaluated in multiple environments and
several genotypes were identified with yield comparable to that of high
yielding check cultivars with seed protein and/or oil superior to the
checks. These genotypes will provide breeders with additional sources of
germplasm for continuing efforts to improve seed composition traits
without compromising seed yield, and provide growers with more
profitable varieties.}

@@

Seed yield, oil, and protein are all valuable traits in a soybean
variety, however breeding lines which have both high yield and high
protein has been difficult to develop due to the negative correlation
between the two traits
\citep{burton1987quantitative, ProtOilCorr, ProtOilCorr_new}. While
considerable efforts have been made to identify loci which control these
seed quality traits so that MAS breeding strategies can be utilized for
their improvement, to date the applications of such markers have been
few. This is largely due to the lack of markers which are uniquely
associated with one trait, and are also stable across genetic and
environmental backgrounds. While there is still reason to continue this
genetic research, it is important that breeders take every opportunity
to identify lines with both high yield, and desirable seed composition
traits like oil and protein content so that new value-added varieties
can be more profitable.

Soybean lines typically contain about 20\% oil and 40\% protein content
on a dry weight basis \citep{ProteinGenomics}. The market for soybean
meal requires 47.5\% protein content in the meal, which corresponds to
approximately 41.5\% protein content with 21\% oil, on a dry weight
basis\citep{ProteinGenomics}. Oil and protein content are two of the
most important seed composition traits in soybean, so if one is
decreased, the other should be correspondingly increased. The inverse
correlation between protein and oil contents is well known and is
suspected to be due at least partially to the action of pleiotropic
genes and competing metabolic pathways which control the expression of
each trait\citep{SeedCompositionGenomics}.

Despite the difficulty in simultaneously breeding for all three of these
traits, varieties with elevated protein contents with little or no yield
reduction has been released by public sector breeders. The high protein
germplasm lines R05-1415 and R05-1772 were released recently that
contain 46.9\% and 46.1\% protein content with 94\% and 91\% yield,
respectively, of the high yielding cultivar
5002T\citep{HiProtHiYield02}. Cultivars TN03-350 and TN04-5321 contain
43.9\% and 43.1\% protein while having superior or comparable yields
with checks cultivars\citep{HiProtHiYield04}.). The Highpro1 cultivar
was released in 2016 and has a yield which is greater than or equal to
that of the check cultivars, while its protein content was 5-6\% higher
than the checks\citep{HighPro1}.Cultivars produced through conventional
breeding techniques such as these have shown that it is possible to
identify lines with both high seed protein and seed yield. Efforts to
find these lines should continue to provide growers and breeders with
additional high value cultivars and germplasm which can be used to
further improve protein and yield traits.

To meet this goal, two recombinant inbred line (RIL) mapping populations
were screened for lines which showed promising combinations of yield and
seed composition traits. Successive rounds of selection were conducted
between 2018 and 2021 to identify and characterize lines with high
values for yield as well as protein and oil composition.

@@

In 2018, oil mapping populations 201 and 202 were grown as plant rows at
the Central Crops Research Station in Clayton, NC. These populations
consisted of 273 and 237 recombinant inbred lines (RILs) respectively.

@@

Following harvest, yield, seed weight, protein, and oil content were
measured after seed was air dried to approximately 7\% moisture content
in a greenhouse. Protein and oil contents were measured on a dry basis
using a Perten DA 7250 NIR®instrument. Yield and seed weight were
measured after seed had been sifted and cleaned of debris.

To select lines for the yield tests in 2019 growing season, lines with
low seed yield or extreme maturity dates in 2018 were removed from
consideration. Two yield trials were then designed for each mapping
population based on maturity dates and desirable agronomic traits,
protein and oil contents. Eighty unique lines were selected from each
population which satisfied these criteria, and each yield test had 40
RILs. Three high-yielding check cultivars and the two parents of the
respective population were also included in each test based on the
maturity spread of the RILs. These yield tests were named test 1 and
test 2 for RILs derived from mapping population 201 and test 3 and 4 for
RILs derived from mapping population 202. Yield check cultivars Dunphy,
Osage, and Roy were used in tests 1 and 2, while Dunphy, Dilday, and
NC-Raleigh were used for tests 3 and 4. The parents for tests 1 and 2
were cultivars LMN09-119 and N09-09, and the parents for tests 3 and 4
were LMN09-19 and N13-47.

@@

A strong negative correlation (r = -0.93) was observed in both tests
between seed protein and seed oil content. A weak negative correlation
(r = -0.29) was observed between seed protein and yield in both tests
which was statistically insignificant in either population. Weak
positive correlations were also observed between seed oil and seed yield
in both populations. A correlation coefficient of r = 0.24 was observed
in Test 1 while a coefficient of r = 0.28 was observed in Test 2. These
correlation coefficients were not statistically significant in either
population. The pairwise correlation coefficients for Tests 1 and 2 are
presented in supplementary figures 2 and 3, respectively.

\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{C:/Users/jhgille2/Documents/R/projects/Jay_yield/exports/plots/lsmean_scatterplot.pdf}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}

@@

Several RILs in each of the tests had yield that was comparable to that
of their check cultivars as per contrasts between RIL means and the
average of the check cultivars for each test. No genotypes had yield
that was higher than that of the yield checks, however many had
comparable yield. Many of these genotypes with comparable yield also had
protein content that was superior to the check average in each test. No
genotypes had oil that was superior to the check cultivars, but some had
comparable yield and oil content as well as superior protein content.
Such genotypes are extremely valuable for production of soybean meal
with higher protein. As mentioned earlier, the meal protein of current
U.S. cultivars is mostly below the minimum market standard of 47.5\%.
The meal protein is calculated by a formula that takes both seed protein
and oil contents into account (reference from Pantalone). The meal
protein contents of these lines with high protein and normal oil
contents ranged from xx -- xx\% compared to the meal protein contents of
the checks (Supplementary table XX).

RILs were also evaluated for lodging and seed quality (data not shown?).
All RILs had seed quality that was on par with that of the high-yielding
checks however, however 11 genotypes had lodging scores that were
statistically significantly greater than the checks on the basis of
contrasts. These RILs were removed from consideration for recommendation
on the basis of their relatively poor agronomic performance
